\thesisabstract{
  This thesis presents a series of publications that contribute novel approaches
  to human neural rendering from a partial information.
  We focus on recently introduced learnable from data radiance fields, namely,
  Neural Radiance Fields (NeRFs) and 3D Gaussian Splatting (3DGS).
  We analyze how these models build an underlying 3D representation from 2D
  images, and ways that the representation can be conditioned to produce
  high-quality renderings of humans.
  We specifically propose several new methods that enable controlling neural
  radiance fields outputs with simple and interpretable inputs from a partial
  training information, and extend this approach to work in a few-shot learning
  setting.

  In the first part of the thesis, we give an overview of the field of neural
  radiance fields.
  We discuss the limitations of the existing approaches and introduce our
  contribution to controllable neural radiance fields.
  Our approach incorportes partial, sparse information given at the training
  time and leverages smoothness of neural networks to enable high-quality and
  controllable human images.

  We then move towards discussing the core limitation of controllable neural
  radiance fields which rely on the availability of high quality data
  annotations for multi-view videos.
  We propose a novel approach to train neural radiance fields in a few-shot,
  multi-view regime by learning an internal representation of subject's
  deformation templates which are mixed smoothly at the inference time.
  We show that such an approach significantly improves the image quality metrics
  over the chosen baselines and opens up new possibilities for human rendering
  from a few images.

  Finally, we show that those approaches can be imbued with an ability to adapt
  their computational requirements at inference time while training the model
  once.
  We introduce a new fine-to-coarse strategy to leraning 3D Gaussian Splatting
  which upsamples a latent 2D grid storing Gaussians' latent representation.
  Our approach achieves comparable results while being deployable to any
  comptation device with a minimal loss of quality.
}

\thesiskeywords{Neural Rendering, Neural Radiance Fields, Few-Shot Learning, Human Rendering, Partial Information}

