\section{Abstract}
  \label{sec:clog-abstract}
  Gaussian Splatting has become popular thanks to its ability to render
  high-quality novel views from casually captured videos efficiently in real
  time.
  However, to achieve high-quality renderings, many Gaussians are used, which
  limits the devices that can render these representations.
  In this work, we introduce a method to dynamically change the number of
  Gaussians used on demand, according to where the representation is deployed
  and how many Gaussians can be supported.
  Importantly, we achieve this with a single representation that is trained
  only once---a representation that embeds a continuous notion of levels of
  detail.
  Specifically, we train a UV mapping---a multi-channel 2D image---of 3D
  Gaussian features, which is then interpreted by a small Multi-Layer
  Perceptron (MLP) to the characteristics of the Gaussian, i.e., center,
  covariance, opacity, and color.
  When training, we make sure that downsampled versions of this image also
  render the scene well by introducing a modulator that adjusts UV map
  features after downsampling.
  Thus, at inference time, this 2D image can be downsampled to any resolution
  of choice, rendering any number of 3D Gaussians.
  We demonstrate the performance of our method on
  NeRF~\cite{mildenhall2020nerf}, DTU~\cite{aanaes2016large} and
  AVA~\cite{martinez2024codec} at with various number of Gaussians,
  outperforming the state-of-the-art 3D Gaussian Splatting representations
  that support levels of detail.
