\thesisabstract{
  This thesis is a series of publications that introduce novel methods for human neural rendering using limited information, focusing on Neural Radiance Fields (NeRFs) and 3D Gaussian Splatting (3DGS).
  It explores how these models construct 3D representations from 2D images and
  demonstrates ways to condition these representations for generating
  high-quality human renderings.
  We propose techniques that use simple, interpretable inputs derived from
  sparse training data and extends these methods to perform effectively in
  few-shot learning scenarios.

  We begin by examining the field of neural radiance fields, addressing
  limitations in existing approaches and presenting contributions to
  controllable radiance fields.
  By incorporating partial and sparse data during training, it leverages the
  smoothness of neural networks to produce controllable, high-quality human
  images.

  To tackle the reliance on extensive, high-quality data annotations from
  multi-view videos, we introduce a new method for training neural radiance
  fields in few-shot, multi-view settings.
  This approach learns internal deformation templates, which blend smoothly
  during inference, significantly improving image quality compared to existing
  baselines and enabling effective human rendering from limited input images.

  The work also addresses the need for adaptable computational efficiency during
  inference.
  It proposes a fine-to-coarse learning strategy for 3D Gaussian Splatting,
  which upscales a latent 2D grid that stores Gaussian representations.
  This strategy achieves competitive results while allowing deployment on
  various computational devices with minimal quality loss.

  In addition, we develop a novel model for controlling radiance fields through
  environmental lighting.
  By incorporating precomputed radiance transfer, this model enables physically
  plausible scene relighting and provides users with intuitive control over
  lighting in reconstructed scenes.

  This research advances the state of the art in controllable neural radiance
  fields and expands their application to few-shot learning scenarios.
  These innovations enhance the possibilities for human rendering from limited
  information and open new directions for future research in the field.
}

\thesiskeywords{Neural Rendering, Neural Radiance Fields, Few-Shot Learning, Human Rendering, Partial Information, Gaussian Splatting}

